Examining the Influence of Premises Type and Time of Day on Violent Crime in Toronto

### 2024-12-03 paper/paper.qmd
minor updates

We use the statistical programming language R [@citeR] to process, clean, and analyze the dataset for this study. Key R packages employed include `tidyverse` for data manipulation and visualization [@thereferencecanbewhatever], `arrow` for efficient data storage and retrieval [@arrowR], `here` for managing file paths [@here], `rstanarm` for Bayesian modeling [@rstanarm], `kableExtra` for enhanced table formatting [@kableExtra], `janitor` for cleaning messy data [@janitor], `lubridate` for working with dates and times [@lubridate], `marginaleffects` for calculating marginal effects [@marginaleffects], `plumber` for creating REST APIs [@plumber], `testthat` for unit testing [@testthat], and `opendatatoronto` for accessing Toronto’s open data [@opendatatoronto].

Our data is sourced from the Open Data Toronto portal [@OpenDataToronto], specifically from the "Motor Vehicle Collisions Involving Killed or Seriously Injured Persons" dataset, which captures detailed information on traffic incidents in Toronto where individuals were either killed or seriously injured. This dataset, compiled by the Toronto Police Service, includes variables that record the time, location, environmental conditions, and specific parties involved in each collision, showing factors that may increase the risk of severe outcomes.

The detailed information within the dataset enables a rigorous analysis of patterns in collision severity. By examining factors such as driver condition, road surface, and lighting conditions, we aim to identify high-risk scenarios that may require targeted safety interventions. Following principles outlined in "Telling Stories with Data" [@tellingstories], this study prioritizes transparency and reproducibility in presenting findings that can inform road safety policies.

@@

Figure 7 (see Section B.3) shows the range of coefficient estimates for our model within the 90% probability interval. However, because the credibility intervals for several predictors are relatively narrow, particularly for premises types like "Transit," it is challenging to observe the trends clearly. To address this, Figure 8 was created with the x-axis limited to a range of -5 to 5. Combining Figures 7 and 8, we observe statistical significance for several predictors, including "Transit," "Afternoon," and "Evening," as their credibility intervals do not cross zero.

@@

Figures @fig-modelresults1 and @fig-modelresults2 visualize the 90% credible intervals for the predictors in the Bayesian logistic regression model. These plots provide intuitions into the uncertainty surrounding the estimated coefficients for each predictor.

Figure @fig-modelresults1 displays the 90% credible intervals for all predictors without restrictions on the x-axis. This plot shows how the predictors such as `PREMISES_TYPE` and `TIME_OF_DAY` influence the likelihood of violent crimes. For example, `PREMISES_TYPETransit` has a high positive credible interval, indicating a significant increase in the likelihood of violent crime in transit areas. Conversely, `PREMISES_TYPECommercial` and `PREMISES_TYPEEducational` have negative credible intervals, indicating a lower likelihood of violent crimes in these premises types.

In Figure @fig-modelresults2, the x-axis has been restricted to a range of -5 to 5 to better visualize predictors with smaller intervals. This refined view helps highlight subtle but meaningful differences between categories like `TIME_OF_DAYAfternoon` and `TIME_OF_DAYEvening`, which show a noticeable increase in the likelihood of violent crimes compared to the reference category (`TIME_OF_DAYEarly Morning`).

>>>>

We use the statistical programming language R [@citeR] to process, clean, and analyze the dataset for this study. Key R packages employed include `tidyverse` for data manipulation and visualization [@thereferencecanbewhatever], `arrow` for efficient data storage and retrieval [@arrowR], `here` for managing file paths [@here], `rstanarm` for Bayesian modeling [@rstanarm], `kableExtra` for enhanced table formatting [@kableExtra], `janitor` for cleaning messy data [@janitor], `lubridate` for working with dates and times [@lubridate], `marginaleffects` for calculating marginal effects [@marginaleffects], `plumber` for creating REST APIs [@plumber], `testthat` for unit testing [@testthat], and `opendatatoronto` for accessing Toronto’s open data [@OpenDataToronto].

Our data is sourced from the Open Data Toronto portal [@OpenDataToronto], specifically from the "Major Crime Indicators" dataset, which records reported crimes across Toronto. Compiled by the Toronto Police Service, the dataset includes details such as crime categories, locations, and times, enabling an analysis of crime trends and patterns.

By examining factors like crime type, neighborhood, and temporal distribution, this study aims to identify areas and times of higher crime occurrence. The findings can support targeted interventions and resource allocation. Following principles outlined in Telling Stories with Data [@tellingstories], this study prioritizes clarity and reproducibility to inform community safety strategies effectively.

@@

@fig-modelresults1 shows the range of coefficient estimates for our model within the 90% probability interval. However, because the credibility intervals for several predictors are relatively narrow, particularly for premises types like "Transit," it is challenging to observe the trends clearly. To address this, @fig-modelresults2 was created with the x-axis limited to a range of -5 to 5. Combining @fig-modelresults1 and @fig-modelresults2, we observe statistical significance for several predictors, including "Transit," "Afternoon," and "Evening," as their credibility intervals do not cross zero.

@@

@fig-modelresults1 and @fig-modelresults2 visualize the 90% credible intervals for the predictors in the Bayesian logistic regression model. These plots provide intuitions into the uncertainty surrounding the estimated coefficients for each predictor.

@fig-modelresults1 displays the 90% credible intervals for all predictors without restrictions on the x-axis. This plot shows how the predictors such as `PREMISES_TYPE` and `TIME_OF_DAY` influence the likelihood of violent crimes. For example, `PREMISES_TYPETransit` has a high positive credible interval, indicating a significant increase in the likelihood of violent crime in transit areas. Conversely, `PREMISES_TYPECommercial` and `PREMISES_TYPEEducational` have negative credible intervals, indicating a lower likelihood of violent crimes in these premises types.

In @fig-modelresults2, the x-axis has been restricted to a range of -5 to 5 to better visualize predictors with smaller intervals. This refined view helps highlight subtle but meaningful differences between categories like `TIME_OF_DAYAfternoon` and `TIME_OF_DAYEvening`, which show a noticeable increase in the likelihood of violent crimes compared to the reference category (`TIME_OF_DAYEarly Morning`).
